---
title: "Peer Assignment 1"
author: "SMinkes"
date: "Sunday, February 15, 2015"
---

Peer Assignment 1, by SMinkes
===========================================================

Load the data in a suitable format

```{r echo=TRUE}
data <- read.csv("activity.csv", header=TRUE, sep=",")
```

Load the required packages for calculating the total number of steps

```{r echo=TRUE, results="hide"}
install.packages("plyr")
library(plyr)
install.packages("lattice")
library(lattice)
```

Calculate the total number of steps taken per day

```{r echo=TRUE, results="asis"}
num_steps <- ddply(data, c("date"), summarise, sum = sum(steps), mean = mean(steps, na.rm=TRUE))
```

Create a Histogram

```{r Histogram number of steps, echo=TRUE}
hist(num_steps$sum)
```

Calculate the mean and the median

```{r echo=TRUE}
mean(num_steps$sum, na.rm=TRUE)
median(num_steps$sum, na.rm=TRUE)
```

Calculate the 5 min interval averages

```{r echo=TRUE}
five <- ddply(data, c("interval"), summarise, average = mean(steps, na.rm=TRUE))
```

Plot the 5 min intervals
```{r 5min interval plot, echo=TRUE}
plot(five, type="l")
```

Calculate the max

```{r echo=TRUE}
maxrow <- five[which.max(five$average),]
maxrow$interval
```

Calculate the number of missing steps

```{r echo=TRUE}
length(which(is.na(data$steps)))
```

Impute missing values with a zero and save it in a new data frame

```{r echo=TRUE}
new <- replace(data, is.na(data), 0)
```


Create a new plot and calculate the mean and the median
```{r Histogram newly calculated steps, echo=TRUE}
new_steps <- ddply(new, c("date"), summarise, sum = sum(steps))
hist(new_steps$sum)
mean(new_steps$sum)
median(new_steps$sum)
```

Create a function to determine whether a day is in the weekend or is a weekday. Apply the function to each day in the list

```{r echo=TRUE}
weekend <- function(date) {
  day <- weekdays(as.Date(as.character(date), "%Y-%m-%d"))
  if(day == "Sunday" || day == "Saturday") {
    day <- "weekend" 
  } else {
    day <- "weekday"
  }
  as.factor(day)
}

new$day <- sapply(new$date, weekend)
```

Plot the panel plot
```{r Panelplot, echo=TRUE}
panel_plot_data <- ddply(new, c("day", "interval"), summarise, average = mean(steps))
xyplot(average ~ interval | day, data = panel_plot_data, layout = c(1,2), type="l")
```